---
title: "FlightControl"
author: "Joerg Steinkamp"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FlightControl}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Get the raw data

This is only relevant, if you intent to use your own recorded flight tracks. Otherwise you cal jump down to section "Get it into R".

To capture the data, you need a SDR-RTL dongle and the software [dump1090](https://github.com/antirez/dump1090). Compile it and run the command in a terminal shell
```{bash, eval=FALSE}
./dump1090 --aggressive --quiet --net
```
On port 30003 of localhost, you can capture the result via the commandline
```{bash, eval=FALSE}
nc -d 127.0.0.1 30003 >> flights_$(date +\%Y\%m\%d).csv
```

I did this on a [Debian Linux](https://www.debian.org/) desktop and on a [Raspberry Pi](https://www.raspberrypi.org/) with Raspbian.

## Get it into R

Load the package:

```{r}
library(FlightControl)
```

And run `covert` with the recorded dates the path to where the csv files are (`source.dir`) and where to save the RData files (`target.dir`). This converts the raw CSV files above in much smaller RData files. However, allready with a tiny bit of loss, since `convert` calls `airplanes` with its default parameters. You can override this by setting `raw=TRUE` or  changing `min.altitude`/`max.altitude` and/or `unit="feet"`. Setting raw to TRUE returns the data as is. Changing only the min/max altitude matches the flight IDs and flight positions and converts feet to meters, if unit isn't set to "feet". Currently only feet and meter are supported, anything else given as unit is treated as meters. 

```{r, eval=FALSE}
dates <- c(strptime("2016-10-28", format="%Y-%m-%d"),
           strptime("2016-10-30", format="%Y-%m-%d"))
convert(dates, "data", "data", max.altitude=3000)
```

Now import those .RData files and append them if desired by passing either one date or several to `import`. And since several callSigns appear several times a day, e.g. when a plane lands and then starts again, those have to be split, which is done by `split.pause.callSign` after `pause.limit` seconds (default 900).

```{r, eval=FALSE}
flights <- import(dates, source.dir = "data", verbose=FALSE)
flights <- split.pause.callSign(flights)
```

Or as done here use the built in dataset.

```{r}
data(flights)
flights <- split.pause.callSign(flights)
```

Apply some filters to delete unrealistic coordinates and too short flight tracks (less than 10 nodes). 

```{r}
flights <- filter.position(flights, west=0, east=15, south=45, north=55,length = 10)
```

## Visualization

This is not part of the package, since it depends on a lot of other packages, which do not necessarily need to be installed. Therefore, the following code is not executed as part of this vignette, just listed as example code!

### Flat 2D maps

Below  is a function to create a map with the flight-tracks above an [OpenStreetMap](https://www.openstreetmap.org). The R package `OpenStreetMap` depends on rJava, which I experienced some trouble with in configuring. Therefore I pulled that dependecy out of my FlightControl package and list it here as example.
```{r}
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(OpenStreetMap))
suppressPackageStartupMessages(require(viridis))
suppressPackageStartupMessages(require(raster))
#' Map of recorded flight tracks
#'
#' Creates a map of flight tracks with an underlying OpenStreetMap layer.
#'
#' @param flights the flights data.frame
#' @return ggplot2 object
#' @export
#' @author Joerg Steinkamp \email{joergsteinkamp@@yahoo.de}
#' @import OpenStreetMap
#' @importFrom ggplot2 autoplot aes geom_path annotate scale_colour_gradientn xlab ylab
#' @importFrom viridis inferno
#' @importFrom raster extent
map.flights <- function(flights) {
  lon=lat=altitude=callSign=NULL
  extent <- extent(x=c(min(flights$lon), max(flights$lon)),
                   y=c(min(flights$lat), max(flights$lat)))
  extent@xmin <- extent@xmin - (extent@xmax-extent@xmin) * 0.05
  extent@xmax <- extent@xmax + (extent@xmax-extent@xmin) * 0.05
  extent@ymin <- extent@ymin - (extent@ymax-extent@ymin) * 0.05
  extent@ymax <- extent@ymax + (extent@ymax-extent@ymin) * 0.05

  ## retrieve the OpenSteetMap data and reproject it
  map <- openmap(c(extent@ymax, extent@xmin), c(extent@ymin, extent@xmax),
                 type = "apple-iphoto", minNumTiles=10)
  mapLatLon <- openproj(map)

  p <- autoplot(mapLatLon)
  p <- p + geom_path(data=flights,
                     aes(x=lon, y=lat, color=altitude, group=callSign),
                     alpha=0.5, size=0.8)
  p <- p + scale_colour_gradientn(colours=inferno(255))
  p <- p + xlab("longitude") + ylab("latitude")
  return(p)
}
```

Then create a map of one day with ascending (2016-10-28) and one day with descending (2016-10-30) airplanes above Mainz with the above function.

```{r, fig.width=7, fig.height=5}
flights.asc <- filter.direction(subset(flights, date < as.POSIXlt("2016-10-28 24:00:00")), direction="asc")
flights.dec <- filter.direction(subset(flights, date > as.POSIXlt("2016-10-30 00:00:00")), direction="dec")
map <- map.flights(flights.asc)
map + labs(title="Ascending (2016-10-28)")
map <- map.flights(flights.dec)
map + labs(title="Decending (2016-10-30)")
dev.off()
```

### 3D with rgl

For the terrain I prepared a digital elevation model (`dem`) based on the [Shuttle Radar Topographic Mission (SRTM)](http://www2.jpl.nasa.gov/srtm/), which is needed as ground surface. If you want to use  the package for another region, you need to get and prepare your own DEM data.

First filter out those flights not to include in the analysis, e.g. remove those where the callSign starts with 'DM' (local flight tracks at low altitude).
```{r, eval=FALSE}
flights <- subset(flights, !grepl("^DM", callSign))
```

There are two possible OpenGL 3D plots, one with individual flight tracks (takes very long) and one with a transparent 'carpet' of the minimum flight altitude remapped on a 1km grid (only a bit faster).

```{r, eval=FALSE}
##animate.flights.3d(flights, dem)
```

For the transparent carpet several days should be used to achieve a smoother surface, which is done in the following example (Oct. 24 - 30, 2016). And several more filters had to be applied: e.g. a few logged position were in Asia or America, which can not be true.

```{r, eval=FALSE}
##animate.flights.3d(flights, dem, tracks = FALSE)
```

For the video the same as above for the 2D movie is valid. If the commandline tool 'ffmpeg' is not available the pictures ar in R's temporyry folder. 